Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
merge

\begin{equation}
S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
\end{equation}

@@

From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + k(t+ Qt)$.  Of this, the only parallelizable part is the $kQt$ term.  Thus to find the expression of the non-parallelizable portion, we simplify:

\[
B = \frac{t+kt}{t+kt+kQt} = \frac{1+k}{kQ+k+1}
\]

@@

\[
\begin{aligned}
S(n) &= \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{1}{\left( B + \frac{1}{n} (1-B) \right)}\\
       &= \frac{T(1)}{T(1) \left( \frac{1+k}{kQ+k+1} + \frac{1}{n} (1- \frac{1+k}{kQ+k+1}) \right)} \\
       &= \frac{n(kQ+k+1)}{n+nk+kQ}
\end{aligned}
\]

As written below, we have used $Q = 100$ with various $k$.

When $k = 2$ and $Q = 100$, for maximal speedup (fewest iterations) of $S(n) = \frac{203n}{3n+200} \approx 67$

When $k = 10$ and $Q = 100$, for more iterations and less error, we'd see an Amdahl bounded max speedup of $S(n) = \frac{1011n}{11n+1000} \approx 91$

It is important to note that this speedup is measured with respect to the parareal agorithm done in parallel. This does not compare the Parareal to doing $g_{\textnormal{fine}}$ but rather shows how given infinite processors, our speedup will approach certain asymptotic bounds. We do see however large constants in the denominator which means that we will get better speedup as we increase n past a certain size.

@@

\[
S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
\]

@@

\begin{center}
\begin{tabular}{|c|c|c|c|c|} \hline
($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
(128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
(128, 128) & 4 & 2 & 2.31 \\ \hline
(128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
(128, 128) & 8 & 8 & (still queued on Odyssey) \\ \hline
(256, 256) & 2 & 2 & 2.57 \\ \hline %real
(256, 256) & 4 & 2 & 11.32 \\ \hline
(256, 256) & 4 & 4 & 13.65 \\ \hline
(256, 256) & 8 & 8 & 17.44 \\ \hline
(512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
(512, 512) & 4 & 2 & (still queued on Odyssey)\\ \hline
(512, 512) & 4 & 4 & (still queued on Odyssey)\\ \hline %check
(512, 512) & 8 & 8 & (still queued on Odyssey) \\ \hline %check
\end{tabular}
\end{center}

@@

@@

\subsection{Demonstrating Scaling and Efficiency}

\subsection{Parallel in Space}

bsim TODO eventually

@@

TODO NUMPY WRITE FATST HING

@@

TODO WRITE ABOUT BETTER DATA COLLECTION

>>>>

\begin{equation}
S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
\end{equation}

@@

From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + K(t+ Qt)$.  Of this, the only parallelizable part is the $kQt$ term.  Thus to find the expression of the parallelizable portion, we simplify:

\[
AA
\]

@@

\[
S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
\]

@@

\[
S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
\]

@@

\begin{center}
\begin{tabular}{|c|c|c|c|c|} \hline
($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
(128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
(128, 128) & 4 & 2 & 2.31 \\ \hline %check
(128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
(128, 128) & 8 & 8 & (still queued on Odyssey) \\ \hline
(256, 256) & 2 & 2 & 2.57 \\ \hline %real
(256, 256) & 4 & 2 & 11.32 \\ \hline %check
(256, 256) & 4 & 4 & 13.65 \\ \hline %check
(256, 256) & 8 & 8 & 17.44 \\ \hline %check
(512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
(512, 512) & 4 & 2 & (still queued on Odyssey)\\ \hline
(512, 512) & 4 & 4 & (still queued on Odyssey)\\ \hline %check
(512, 512) & 8 & 8 & (still queued on Odyssey) \\ \hline %check
\end{tabular}
\end{center}

@@

\subsection{Parallelism Paradigm Comparisons}

Now with 3 different parallel implementations, two in time with the parareal, one in space, we compare across the methods to see if we can make any statements about general performance of our different paradigms.

TODO WESLEY

TODO compare the computation times (and perhaps a slight change in accuracy of the results) between both parareal methods the best in the sections before hand

TODO bsim and ashi show graphs and some times for the forward Euler parareal
versions---the fastest ones or both if we have both implemented

\subsection{Comparison to C++ Implementations}

The numpy environment combined with the required structure of MPI required
certain type/data structure conversions which will take time.not not cunot cu

@@

\subsection{Demonstrating Scaling}

ashi TODO by taking some numbers and following

@@

@@

TODO Wesley
